---
title: 'Deliverable #1 Project Proposal'
author: "Group #1: Holland Robinson, Vu Nguyen, Kathryn Mioduszewski, Pranav Patil, Sreechandana Anumolu, Yash Ghate"
date: "4/4/2023"
output: 
  html_document:
    toc: true
    theme: readable
    highlight: tango
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What Data Will You Use?
This dataset details used car data. The Y variable is whether or not the car is sold. The X variables include attributes like car type, mileage, make, and model, among other attributes.
```{r}
car <- read.csv("Car_data.csv")
```

## Business Question
With this data, you can build an algorithm to predict what attributes influence the sale of a car. This is relevant to used car dealerships when choosing whether you would accept this car from the seller to subsequently sell. 

## Preliminary Exploration and Cleaning of Data
```{r}
car$Sales_ID <- NULL
car$age <- max(car$year) - car$year
car$name <- as.factor(car$name)
car$State.or.Province <- as.factor(car$State.or.Province)
car$Region <- as.factor(car$Region)
car$seller_type <- as.factor(car$seller_type)
car$transmission <- as.factor(car$transmission)
car$fuel <- as.factor(car$fuel)
car$owner <- as.factor(car$owner)
car$sold <- as.factor(car$sold)
car$name <- ifelse(car$name == "Opel" | car$name == "Ashok" | car$name == "MG" | car$name == "Daewoo" | car$name == "Kia" | car$name == "Ambassador" | car$name == "Isuzu" | car$name == "Land" | car$name == "Force" , "Other", car$name)
car$City <- ifelse(car$City == "New York City" | car$City == "Los Angeles" | car$City == "Seattle" | car$City == "Chicago" | car$City == "Boston" | car$City == "Washington" | car$City == "Philadelphia" | car$City == "Charlotte" | car$City == "Miami" | car$City == "Detroit"  , car$City, "Other")

car$sold <- as.numeric(car$sold) - 1


str(car)
summary(car)
```

## Preliminary Regression Model
```{r}
#Test and Train

set.seed(12345)
test_rows <- sample(1:nrow(car), 0.25*nrow(car))

car_test <- car[test_rows, ]
car_train <- car[-test_rows, ]

# Model

basemodel <- glm(sold ~ ., data = car_train, family = "binomial")
summary(basemodel)

# Predict
carpred <- predict(basemodel, car_test, type = "response")
summary(carpred)

car_binary_pred <- ifelse(carpred >= 0.5, 1, 0)
table(car_binary_pred)

# Confusion Matrix
library(caret)
car_test$sold <- ifelse(car_test$sold, 1, 0)
confusionMatrix(as.factor(car_binary_pred), as.factor(car_test$sold))
```

## GitHub

[Here](https://github.com/Hmrobinson/to628-group1project) is the link to our GitHub Repo. The Professor and TA's have been added as collaborators. 

