---
title: 'Deliverable #1 Project Proposal'
author: "Group #1: Holland Robinson, Vu Nguyen, Kathryn Mioduszewski, Pranav Patil, Sreechandana Anumolu, Yash Ghate"
date: "4/14/2023"
output: 
  html_document:
    toc: true
    theme: readable
    highlight: tango
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What Data Will You Use?
This dataset details used car data. The Y variable is whether or not the car is sold. The X variables include attributes like car type, mileage, make, and model, among other attributes.

```{r}
car <- read.csv("Car_data (1).csv")
```

## Business Question
With this data, you can build an algorithm to predict what attributes influence the sale of a car. This is relevant to used car dealerships when choosing whether you would accept this car from the seller to subsequently sell. 

## Preliminary Exploration and Cleaning of Data
```{r}
car$Sales_ID <- NULL
car$age <- max(car$year) - car$year
car$name <- as.factor(car$name)
car$State <- as.factor(car$State)
car$Region <- as.factor(car$Region)
car$seller_type <- as.factor(car$seller_type)
car$transmission <- as.factor(car$transmission)
car$fuel <- as.factor(car$fuel)
car$owner <- as.factor(car$owner)
car$sold <- as.factor(car$sold)
car$name <- ifelse(car$name == "Opel" | car$name == "Ashok" | car$name == "MG" | car$name == "Daewoo" | car$name == "Kia" | car$name == "Ambassador" | car$name == "Isuzu" | car$name == "Land" | car$name == "Force","Other", car$name)

car$City <- ifelse(car$City == "New York City" | car$City == "Los Angeles" | car$City == "Seattle" | car$City == "Chicago" | car$City == "Boston" | car$City == "Washington" | car$City == "Philadelphia" | car$City == "Charlotte" | car$City == "Miami" | car$City == "Detroit"  , car$City, "Other")

carv1 <- as.data.frame(model.matrix(~.-1,car))

str(car)
summary(carv1)

```

#Normalize Data 
```{r}
normalize <- function(x){
  return((x-min(x))/(max(x)-min(x)))
}

car_norm <- as.data.frame(lapply(carv1, normalize))

summary(car_norm)

```

#Split Data 
```{r}

set.seed(12345)
test_rows <- sample(1:nrow(car_norm), 0.5*nrow(car_norm))
car_test <- car_norm[test_rows, ]
car_train <- car_norm[-test_rows, ]
cardt_test<-car[test_rows,]
cardt_train<-car[-test_rows,]

summary(car_train)

```

#Regression Model
```{r}

# %%%%%  MODEL %%%%%% 

basemodel <- glm(soldY ~., data = car_train, family = "binomial")

summary(basemodel)

# %%%%% Predict %%%%%

library(caret)
carpred <- predict(basemodel, car_test, type = "response")
car_binary_pred <- ifelse(carpred >= 0.5, 1, 0)
table(car_binary_pred)
car_test$sold <- ifelse(car_test$sold, 1, 0)
confusionMatrix(as.factor(car_binary_pred), as.factor(car_test$sold))

```


# KNN Model
```{r}

#Data Cleaning & creating test and train 
carknn_test <- car_norm[test_rows,-match("soldY",names(car_norm))]
carknn_train <- car_norm[-test_rows,-match("soldY",names(car_norm))]
carknn_test_labels <- car_norm[test_rows,"soldY"]
carknn_train_labels <- car_norm[-test_rows,"soldY"]

#Build Model
library(class)
knn_model <- knn(carknn_train,carknn_test,carknn_train_labels,k=40,prob= TRUE)

## Evaluate Model
library(caret)
confusionMatrix(as.factor(knn_model), as.factor(carknn_test_labels),positive = "1")

```

# ANN Model
```{r}
library(neuralnet)

#BUILDING THE MODEL

model_ann1 <- neuralnet(soldY ~., data = car_train, hidden = 5,2) 
plot(model_ann1)
saveRDS(model_ann1, file = 'model_ann1.rds')

# Predicting the ANN model 
model_ann_1predict <- predict(model_ann1,car_test)
model_ann1bin <- ifelse(model_ann_1predict >= 0.25,1,0)

#EVALUATING 
library(caret)
confusionMatrix(as.factor(model_ann1bin), as.factor(car_test$soldY), positive = "1")

```

# DECISION TREE
```{r}

library(C50)

#BUILDING THE MODEL
set.seed(12345)
cartree <- C5.0(as.factor(sold)~., data = cardt_train)
plot(cartree)
summary(cartree)

#PREDICTING THE MODEL
cartreepred <- predict(cartree, cardt_test)

#EVALUATING THE MODEL
confusionMatrix(as.factor(cartreepred),as.factor(cardt_test$sold), positive = "Y")

```

# RANDOM FOREST
```{r}
library(randomForest)
#Predict the RF Model 
carforest <- randomForest(as.factor(soldY) ~. , data = car_train)
summary(carforest)
randomForest::varImpPlot(carforest)

#Evaluate and Predict 
carrfpredict <- predict(carforest, car_test)
summary(carrfpredict)
confusionMatrix(as.factor(carrfpredict), as.factor(car_test$soldY), positive = "1")

```

#Stacked Model 
```{r}

stacked_car <- data.frame(log_pred = carpred, KNN = knn_model, ANN = model_ann_1predict, DT_pred = cartreepred, rf_pred = carrfpredict,y = car_test$soldY)

summary(stacked_car)

```

#Test and Train
```{r}

set.seed(1234)
test_set1 <- sample(1:nrow(stacked_car), 0.3*nrow(stacked_car)) 
cars_train <- stacked_car[-test_set1, ]
cars_test <- stacked_car[test_set1, ]

```

# Building Stacked Model's Decision tree

```{r}

library(C50)
stackedtree <- C5.0(as.factor(y) ~., data = cars_train)
plot(stackedtree)

```

#Predict and Evaluate Model 
```{r}
library(caret)
stackedpredict <- predict(stackedtree,cars_test)
confusionMatrix(as.factor(stackedpredict), as.factor(cars_test$y), positive = "1")

```

# Cost Matrix 
```{r}
error_cost = matrix(c(0,1,4,0), nrow =2) 
error_cost

errormodel <- C5.0(as.factor(soldY)~., data = car_train,costs = error_cost)
prederror <- predict(errormodel, car_test)
summary(prederror)
confusionMatrix(as.factor(prederror), as.factor(car_test$soldY), positive = "1")

```

